---
title: "p8105_hw3_yl5505"
author: "Yan Li"
date: "2024-10-11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(knitr)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
library(p8105.datasets)
data("ny_noaa")
```
# Problem 1

## Overview

```{r}

summarize(ny_noaa, rows = n(), columns = ncol(ny_noaa))

summarize(ny_noaa, across(everything(), ~mean(is.na(.))))
```

-- This dataset is a weather dataset, it has 2595176 rows and 7 variables
- `id`: Weather station ID
- `date`: Date of observation
- `prcp`: Precipitation (tenths of mm)
- `snow`: Snowfall (mm)
- `snwd`: Snow depth (mm)
- `tmax`: Maximum temperature (tenths of degrees C)
- `tmin`: Minimum temperature (tenths of degrees C)
-- There are many columns with missing data. tmax and tmin have about 44% data missing, snow has 15%, and snwd has 23%. Missing data could cause bias or make the dataset less reliable for certain analyses. 

## Data cleaning
- create new variables
```{r}
ny_noaa %>% 
  count(snow) %>%
  arrange(desc(n))

ny_noaa = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin))
```

-- The most commonly observed snowfall value is 0 mm, it mean that on most days there was no snowfall recorded. It is reasonable becasue New York does not snow a lot considering its location and geography. 

## Two-panel plot
- my codes won't run :(
```{r}
ny_noaa %>% 
  group_by(id, year, month) %>% 
  filter(month %in% c(1, 7)) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE, color = id)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id)) + geom_point() + geom_path() +
  facet_grid(~month) +
  labs(title = "Mean monthly temperature for each station across years for January and July")
```

## tmax vs tmin
- with some help
```{r}

# (i) Tmax vs Tmin Plot
plot_tmax_vs_tmin = ny_noaa %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex(bins = 50) +
  scale_fill_viridis_c() +
  labs(
    title = "Hexbin Plot of Tmax vs Tmin",
    x = "Minimum Temperature (°C)",
    y = "Maximum Temperature (°C)",
    fill = "Count"
  )

# (ii) Distribution of Snowfall (>0 and <100) by Year
plot_snowfall_distribution = ny_noaa %>%
  filter(snow > 0, snow < 100) %>%
  ggplot(aes(x = snow, y = factor(year), fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis_c(option = "C") +
  labs(
    title = "Distribution of Snowfall (0 < Snow < 100 mm) by Year",
    x = "Snowfall (mm)",
    y = "Year",
    fill = "Snowfall (mm)"
  ) 

# Combine the two plots
combined_plot = plot_tmax_vs_tmin + plot_snowfall_distribution + plot_layout(ncol = 1)

# Display the combined plot
combined_plot

```




#Problem 2
```{r}
hw_covar = read_csv("nhanes_covar.csv",skip = 4 )|>
  janitor::clean_names()

hw_accel = read_csv("nhanes_accel.csv")|>
  janitor::clean_names()

```
- merge two dataset
```{r}
hw_ac = inner_join(hw_accel, hw_covar, by = "seqn")
hw_ac
```

```{r}
ac_clean = hw_ac |>
  filter(age >= 21) |>
  drop_na()|>
  mutate(
    sex = factor(sex, levels = c(1, 2), labels = c("male", "female")),
    education = factor(education, levels = c(1, 2, 3), labels = c("Less than high school", "High school equivalent", "More than high school"))
  )
```




## create table for the number of men and women in each education category
```{r}
edu_table = ac_clean |>
  group_by(education, sex) |>
  summarize(count = n(), .groups = "drop") |>
  pivot_wider(names_from = sex, values_from = count, values_fill = 0)

edu_table|>kable()
```

## create plot
```{r}
ggplot(ac_clean, aes(x = education, y = age, fill = sex)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("male" = "#1f77b4", "female" = "#ff7f0e")) +
  labs(
    title = "Age Distribution by Education Level and Gender",
    x = "Education Level",
    y = "Age",
    fill = "Gender"
  )
```
- The table shows the gender distribution across education levels,and the "High school equivalent" group has fewer females compared to males.
- The boxplot shows that, in "More than high school", the age distribution is more concentrated, with a visible peak in the 40-45 age range for both men and women.
- Greater variability in age for males across all education levels.

## Total activity Plot

```{r}
ac_clean |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",                
    names_prefix = "min",
    values_to = "activity_value"        
  ) |>
  group_by(seqn, sex, age, education) |>
  summarize(total_activity = sum(activity_value, na.rm = TRUE)) |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +            
  geom_smooth(se = FALSE) +   
  facet_wrap(~ education) + 
  labs(
    title = "Total Daily Activity vs Age by Education Level and Gender",
    x = "Age",
    y = "Total Daily Activity"
  )

```
-The plot shows that physical activity declines with age regardless of education level.
-The trend of higher activity levels among those with more education may indicate that education contributes to increased health awareness.
-In general, women (yellow line) show slightly higher peaks in activity in the younger age groups compared to men.


## 24-hour activity

```{r}
ac_clean |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "activity_value"
  ) |>
  mutate(minute = as.numeric(minute)) |>
  group_by(minute, sex, education) |>
  summarize(mean_activity = mean(activity_value, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = minute, y = mean_activity, color = sex)) +
  geom_line(size = 1, alpha = 0.7) + 
  facet_wrap(~ education) +        
  scale_x_continuous(               
    breaks = seq(0, 1440, by = 240), 
    labels = c("Midnight", "4 AM", "8 AM", "Noon", "4 PM", "8 PM", "Midnight")
  ) +
  labs(
    title = "24-Hour Activity Time Courses by Education Level and Gender",
    x = "Minute of the Day",
    y = "Mean Activity Level",
    color = "Gender"
  )
```
-Both genders have similar low activity levels during nighttime hours, likely reflecting sleep.
-Compare to males, females show consistently higher activity levels during the active parts of the day, especially during morning and evening peaks.

# Problem 3

## Import datasets
```{r}
jan2020 =
  read_csv(file = "./citibike/Jan 2020 Citi.csv", na = c(".", "NA", ""))|>
  janitor::clean_names()|>
  mutate(
    year = 2020,
    month = "January"
  )

jan2024 =
  read_csv(file = "./citibike/Jan 2024 Citi.csv", na = c(".", "NA", ""))|>
  janitor::clean_names()|>
  mutate(
    year = 2024,
    month = "January"
  )

july2020 =
  read_csv(file = "./citibike/July 2020 Citi.csv", na = c(".", "NA", ""))|>
  janitor::clean_names()|>
  mutate(
    year = 2020,
    month = "July"
  )

july2024 =
  read_csv(file = "./citibike/July 2024 Citi.csv", na = c(".", "NA", ""))|>
  janitor::clean_names()|>
  mutate(
    year = 2024,
    month = "July"
  )
```
## Combine 4 dataset
```{r}
bike_com = 
  bind_rows(jan2020,july2020,jan2024,july2024)|>
  relocate(ride_id, year,month, weekdays)

```

## Clean dataset
```{r}
bike_tidy = bike_com |>
  mutate(
    member_casual = as.factor(member_casual),
    weekdays = factor(weekdays, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
    rideable_type= factor(rideable_type),
    month = factor(month, levels = c("January", "July"))
  )
```

## total number of rides in year/month
```{r}
bike_tidy |>
  group_by(year, month, member_casual) |>
  summarize(total_rides = n(), .groups = "drop") |>
  pivot_wider(names_from = member_casual, values_from = total_rides)|>
  knitr::kable()
```
-Members are the primary users of the Citi Bike system.
-July shows much higher ride counts compared to January.
-There is a clear increase in the number of rides between 2020 and 2024 for both casual riders and members.

## 5 most popular starting stations for July 2024
```{r}
bike_tidy |>
  filter(year == 2024, month == "July") |>
  group_by(start_station_name) |>
  summarize(total_rides = n(), .groups = "drop") |>
  arrange(desc(total_rides)) |>
  slice_head(n = 5)|>
  knitr::kable()
```

## the effects of day of the week, month, and year on median ride duration
```{r}
bike_tidy |>
  group_by(year, month, weekdays) |>
  summarize(median_dur = median(duration, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = weekdays, y = median_dur, color = month, shape = month)) +
  geom_point(size = 3, alpha = 0.8) +  
  geom_smooth(aes(group = month),method = "loess", se = FALSE, size = 1) + 
  facet_wrap(~ year) + 
  labs(
    title = "Effects of Day of the Week, Month, and Year on Median Ride Duration",
    x = "Day of the Week",
    y = "Median Ride Duration",
    color = "Month",
    shape = "Month"
  )
```
- Median ride durations are generally higher in July compared to January.
- Ride durations tend to be longer on weekends (Sundays) in both months.
- The patterns are consistent across both 2020 and 2024, but 2024 appears to show a slightly stronger trend.

## impact of month, membership status, and bike type on the distribution of ride duration
```{r}
bike_tidy |>
  filter(year == 2024) |>
  ggplot(aes(x = member_casual, y = duration, fill = rideable_type)) +
  geom_boxplot(alpha = 0.7) + 
  facet_wrap(~ month) + 
  labs(
    title = "Impact of Month, Membership Status, and Bike Type on Ride Duration (2024)",
    x = "Membership Status",
    y = "Ride Duration",
    fill = "Bike Type"
  ) 

```
- Casual riders generally have longer and more varied ride durations compared to members. Ride durations tend to be longer in July compared to January.Electric bikes tend to have longer ride durations compared to classic bikes. 
- There are numerous outliers (especially for casual riders) and these are more prominent for electric bikes in both months.







